---
title: "Web Scraping of Quotes from Famous People using R Take 2"
author: "David Lowe"
date: "November 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

SUMMARY: The purpose of this project is to practice web scraping by gathering specific pieces of information from a website. The web scraping code was written in R and leveraged the rvest package.

INTRODUCTION: A demo website, created by Scrapinghub, lists quotes from famous people. It has many endpoints showing the quotes in different ways, and each endpoint presents a different scraping challenge for practicing web scraping. For this Take2 iteration, the R script attempts to follow the links to the author page and scrape the author information.

Starting URLs: http://quotes.toscrape.com/

## Loading Libraries and Packages

```{r LIBRARY}
library(rvest)
library(jsonlite)
```

## Setting up the necessary parameters

```{r INPUT}
# Specifying the URL of desired web page to be scrapped
starting_url <- 'http://quotes.toscrape.com'

# Creating an html document from the URL
webpage <- read_html(starting_url)
```

## Performing the Scraping and Processing

```{r PAGE OVERVIEW}
dataset_records_css <- html_nodes(webpage, 'div.quote')
head(dataset_records_css)
str(dataset_records_css)
```

```{r ITEM EXTRACTION}
dataset_listing_df <- data.frame(author_name=character(0), author_birthday=character(0), author_birthloc=character(0), author_bio=character(0), author_url=character(0))

done <- FALSE

while (!done) {
  for (item in dataset_records_css) {
  
    author_name_css <- html_nodes(item, 'small.author')
    author_name <- html_text(author_name_css, trim=T)
  
    author_url_css <- html_nodes(item, 'span a')
    author_url_href <- html_attr(author_url_css, 'href')
    author_url <- paste(starting_url, author_url_href, sep = "")
    
    # Avoid duplicating author names
    if (!(author_name %in% dataset_listing_df$author_name)){
      author_page <- read_html(author_url)
      
      author_birthday_css <- html_node(author_page, 'span.author-born-date')
      author_birthday <- html_text(author_birthday_css)
      
      author_birthloc_css <- html_node(author_page, 'span.author-born-location')
      author_birthloc <- html_text(author_birthloc_css)
      
      author_bio_css <- html_node(author_page, 'div.author-description')
      author_bio <- html_text(author_bio_css)
  
      each_record_df <- data.frame(author_name, author_birthday, author_birthloc, author_bio, author_url)
      dataset_listing_df <- rbind(dataset_listing_df, each_record_df)
    }
    
  }
  
  next_page_css <- html_nodes(webpage, 'li.next a')
  if (length(next_page_css)!=0) {
    next_page_url <- html_attr(next_page_css, 'href')
    next_page_url <- paste(starting_url, next_page_url, sep="")
    webpage <- read_html(next_page_url)
    dataset_records_css <- html_nodes(webpage, 'div.quote')
    done <- FALSE
  } else {
    done <- TRUE
  }
  
}
```

## Organizing Data and Producing Outputs

```{r OUTPUT}
cat("Number of author items processed:", nrow(dataset_listing_df))
dataset_listing_json <- toJSON(dataset_listing_df)
write_json(dataset_listing_df, "web-scraping-r-famous-quotes-take2.json")
cat(dataset_listing_json)
```
