---
title: "Web Scraping of Quotes from Famous People using R Take 3"
author: "David Lowe"
date: "November 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

SUMMARY: The purpose of this project is to practice web scraping by gathering specific pieces of information from a website. The web scraping code was written in R and leveraged the rvest package.

INTRODUCTION: A demo website, created by Scrapinghub, lists quotes from famous people. It has many endpoints showing the quotes in different ways, and each endpoint presents a different scraping challenge for practicing web scraping. For this Take3 iteration, the R script attempts to scrape the quote information that is displayed via an infinite scrolling page.

Starting URLs: http://quotes.toscrape.com/scroll

## Loading Libraries and Packages

```{r LIBRARY}
library(rvest)
library(jsonlite)
```

## Setting up the necessary parameters

```{r INPUT}
# Specifying the URL of desired web page to be scrapped
site_url <- 'http://quotes.toscrape.com'
api_url <- 'http://quotes.toscrape.com/api/quotes?page='
page_num <- 1
starting_url <- paste0(api_url, page_num)

# Creating an html document from the URL
webpage <- fromJSON(txt=starting_url)
```

## Performing the Scraping and Processing

```{r PAGE OVERVIEW}
dataset_records_css <- webpage[['quotes']]
head(dataset_records_css)
str(dataset_records_css)
```

```{r ITEM EXTRACTION}
dataset_listing_df <- data.frame(author_name=character(0), quote_text=character(0), quote_tags=character(0), author_url=character(0))

done <- FALSE

while (!done) {
  for (i in 1:nrow(dataset_records_css)) {
  
    item <- dataset_records_css[i,]
    author_name <- item[['author']]$name
    quote_text <- item[['text']]
    quote_tags <- paste(unlist(item[['tags']]), collapse = ", ")
    author_url <- paste0(site_url, item[['author']]['goodreads_link'])
    each_record_df <- data.frame(author_name, quote_text, quote_tags, author_url)
    dataset_listing_df <- rbind(dataset_listing_df, each_record_df)
  }
  
  next_page_exist <- webpage[['has_next']]
  if (next_page_exist) {
    page_num <- webpage[['page']] + 1
    next_page_url <- paste0(api_url, page_num)
    webpage <- fromJSON(txt=next_page_url)
    dataset_records_css <- webpage[['quotes']]
    done <- FALSE
  } else {
    done <- TRUE
  }
  
}
```

## Organizing Data and Producing Outputs

```{r OUTPUT}
cat("Number of author items processed:", nrow(dataset_listing_df))
dataset_listing_json <- toJSON(dataset_listing_df)
write_json(dataset_listing_df, "web-scraping-r-famous-quotes-take3.json")
cat(dataset_listing_json)
```
