---
title: "Web Scraping of Registry of Open Data on AWS Using R"
author: "David Lowe"
date: "October 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

SUMMARY: The purpose of this project is to gather data about the open datasets on AWS. The web scraping code was written in R and leveraged the rvest package.

INTRODUCTION: The Open Data registry exists to help people discover and share datasets that are available via AWS resources. This page lists all usage examples for datasets listed in the registry. [https://registry.opendata.aws/usage-examples]

Starting URLs: https://registry.opendata.aws/

## Loading Libraries and Packages

```{r LIBRARY}
library(rvest)
library(jsonlite)
```

## Setting up the necessary parameters

```{r INPUT}
# Specifying the URL of desired web page to be scrapped
url <- 'https://registry.opendata.aws/'

# Creating an html document from the URL
webpage <- read_html(url)
```

## Performing the Scraping and Processing

```{r PROCESSING}
dataset_records_css <- html_nodes(webpage, '.dataset')
head(dataset_records_css)
str(dataset_records_css)

dataset_listing_df <- data.frame(name=character(0), url=character(0), description=character(0), tags=character(0))

for (item in dataset_records_css) {

  dataset_name_css <- html_nodes(item, 'h3 a')
  dataset_name <- html_text(dataset_name_css)

  detail_url_css <- html_attr(dataset_name_css, 'href')
  detail_url <- paste(url, detail_url_css, sep = "")

  dataset_description_css <- html_nodes(item, 'p')[2]
  dataset_description <- html_text(dataset_description_css)

  dataset_tags_css <- html_nodes(item, 'p span')
  dataset_tags <- character(0)
  for (tag in dataset_tags_css) {
    dataset_tags <- c(dataset_tags, html_text(tag))
  }
  
  dataset_tags <- paste(dataset_tags,collapse=", ")
  each_record_df <- data.frame(dataset_name, detail_url, dataset_description, dataset_tags)
  names(each_record_df) <- c("name", "url", "description", "tags")
  
  dataset_listing_df <- rbind(dataset_listing_df, each_record_df)
}
```

## Organizing Data and Producing Outputs

```{r OUTPUT}
dataset_listing_json <- toJSON(dataset_listing_df)
write_json(dataset_listing_df, "web-scraping-r-aws-opendata-registry.json")
cat(dataset_listing_json)
```
